### 2015-07-31 implementation-concretization.tex
Revisions and better overhead plot.

The next part of the process is iterative.
If any node is a virtual dependency, spack replaces it with with a
suitable interface provider.  It does this by building a reverse
index from virtual packages to providers using the {\tt provides when}
directives (Section~\ref{sec:virtual}). If multiple providers
satisfy the virtual spec's constraints,
Spack consults site and user policies to select the ``best'' possible
provider.  The selected provider may {\it itself} have virtual dependencies,
this process is repeated until there are no more virtual packages in the DAG.

@@

Figure~\ref{fig:concretization-time} shows the running time of concretization vs.
package DAG size in nodes.  We generated this plot by running {\tt concretize}
on all of Spack's 245 packages.  We see that for all but the 9 largest
packages, concretization takes less than 2 seconds on a Sandy Bridge machine, 
and less than 1 second on a Haswell MacBook Pro.  For larger DAGs, we begin
to see superlinear scaling, but even for 50 nodes or more, concretization
takes less than 6 seconds.
%
This performance is expected from a greedy, fixed-point method.
The algorithm repeats, evaluating conditions and adding dependencies and constraints, until
there are no more changes to be made.
%
%  Additionally, our current compiler concretization
% policy requires a search in the DAG; it prefers the compiler of the nearest ancestor
% if the current node's compiler is not set. 
%
More generally, concretization is an instance of the constraint
satisfaction problem, which is NP-complete. Our greedy solver is sufficient for the
packages we have built so far, and its running time is currently negligible. We do
not expect to see software packages with thousands of dependencies in the near
future, and we do not expect concretization to become a bottleneck, even if we
use a full constraint solver.

>>>>

The next part of the process is iterative.
If any node is a virtual dependency, Spack replaces it with with a
suitable interface provider.  It does this by building a reverse
index from virtual packages to providers using the {\tt provides when}
directives (Section~\ref{sec:virtual}). If multiple providers
satisfy the virtual spec's constraints,
Spack consults site and user policies to select the ``best'' possible
provider.  The selected provider may {\it itself} have virtual dependencies,
this process is repeated until there are no more virtual packages in the DAG.

@@

Figure~\ref{fig:concretization-time} shows the running time of concretization vs.
package DAG size in nodes.  We generated this plot by running {\tt concretize}
on all of Spack's 245 packages.  We see that for all but the 9 largest
packages, concretization takes less than 2 seconds on a Sandy Bridge machine, 
and less than 1 second on a Haswell MacBook Pro.  For larger DAGs, we begin
to see a quadratic trend, but even for 50 nodes or more, concretization
takes less than 6 seconds.
%
This performance is expected from a greedy, fixed-point method.
The algorithm repeats, evaluating conditions and adding dependencies and constraints, until
there are no more changes to be made.
%
%  Additionally, our current compiler concretization
% policy requires a search in the DAG; it prefers the compiler of the nearest ancestor
% if the current node's compiler is not set. 
%
More generally, concretization is an instance of the constraint
satisfaction problem, which is NP-complete. Our greedy solver is sufficient for the
packages we have built so far.  Its running time is insignificant
compared to the build time of most packages, and it only executes
once per build.
We do not expect to see software packages with thousands of dependencies in the near
future, and we do not expect concretization to become a bottleneck, even if we
use a full constraint solver.

### 2015-07-31 implementation-environment.tex
Tighten text.
Revisions and better overhead plot.

\subsubsection{Environment isolation}
In addition to inconsistencies that arise from building manually, we
frequently encounter errors due to inconsistencies between the package
installation environment and the environment of the package user.
%
For example, LLNL performance tools use two versions of the {\tt libelf}
library. One is distributed with RedHat Linux, while the
other is publicly available, and they have the same API but an incompatible ABI.
Specifying the wrong version at build time has caused many
unexpected crashes.

Spack manages the build environment by running each call to {\tt install}
in its own process.  In this environment, Spack sets
{\tt PATH}, {\tt PKG\_CONFIG\_PATH}, {\tt CMAKE\_PREFIX\_PATH}, and
{\tt LD\_LIBRARY\_PATH} to include the dependencies of the current build.
These variables are commonly used by build systems to locate dependencies,
and setting them helps to ensure that the correct libraries are detected by
build systems.
The isolated build environment also gives package authors 
free reign to set build-specific environment variables without interfering
with other packages.

\subsubsection{Compiler wrappers and RPATHs}
Finding dependencies at build time is not the only obstacle to reproducible
behavior.  As mentioned in Section~\ref{sec:motivation}, it is also important
for binaries to be able to find dependency libraries at {\it runtime}.
One of the most common user errors at LC is improper library configuration.
Users frequently do not know what libraries a package was built with, and 
it is difficult for them to construct a suitable {\tt LD\_LIBRARY\_PATH} for
a package that was built by someone else.  To reduce the number of support calls,
we typically add {\tt RPATHs} to public software installations, so that paths
to dependencies are embedded in binaries. This way, users do not need to know
details of dependencies to run installed software correctly.

@@

>>>>

\subsubsection{Environment isolation}
In addition to inconsistencies that arise from building manually, we
frequently encounter errors due to inconsistencies between the package
installation environment and the environment of the package user.
%
For example, LLNL performance tools use two versions of the {\tt libelf}
library. One is distributed with RedHat Linux, while the
other is publicly available. They have the same API but an incompatible ABI.
Specifying the wrong version at build time has caused many
unexpected crashes.

Spack manages the build environment by running each call to {\tt install}
in its own process.  It also sets
{\tt PATH}, {\tt PKG\_CONFIG\_PATH}, {\tt CMAKE\_PREFIX\_PATH}, and
{\tt LD\_LIBRARY\_PATH} to include the dependencies of the current build.
Build systems commonly use these variables to locate dependencies,
and setting them helps to ensure that the correct libraries are detected.
Using a separate process also gives package authors 
free reign to set build-specific environment variables without interfering
with other packages.

\subsubsection{Compiler wrappers and RPATHs}
Finding dependencies at build time is not the only obstacle to reproducible
behavior.  As mentioned in Section~\ref{sec:motivation}, it is also important
for binaries to be able to find dependency libraries at {\it runtime}.
One of the most common user errors at LLNL is improper library configuration.
Users frequently do not know what libraries a package was built with, and
it is difficult for them to construct a suitable {\tt LD\_LIBRARY\_PATH} for
a package that was built by someone else.  To reduce the number of support calls,
we typically add {\tt RPATHs} to public software installations, so that paths
to dependencies are embedded in binaries. This way, users do not need to know
details of dependencies to run installed software correctly.

@@

\begin{figure}%{\linewidth}
	\centering
	\includegraphics[width=\columnwidth]{figs/overhead/build-time.pdf}
	\caption{
		Build time on NFS and temp, with and without wrappers.
		\label{fig:build-times}
	}
\end{figure}
\begin{figure}%{\linewidth}
	\centering
	\includegraphics[width=\columnwidth]{figs/overhead/overhead.pdf}
	\caption{
		Build overhead of using NFS and compiler wrappers.
		\label{fig:overhead}`
	}
\end{figure}

\subsubsection{Build Overhead}

Argument parsing and indirection cause Spack's compiler wrappers to incur a small
but noticeable performance overhead.  Figure~\ref{fig:build-time} shows
build times with and without compiler wrappers for seven different packages,
and Figure~\ref{fig:overhead} shows the corresponding overhead as a percentage
of the wrapper-less runtime.  {\tt Dyninst} and {\tt lapack}
use CMake to build, and the rest use autotools.  We ran
each build three times on an Intel Sandy Bridge cluster node, and we
report the average here.
In the case of {\tt dyninst}, the overhead is negligible (-0.4\%),
but it is as high as 12.3\% for shorter builds, like
{\tt mpileaks}

To increase build speed, by default Spack runs each {\tt install}
method in the default temporary directory. On most cluster nodes,
this is a fast, locally-mounted filesystem. In our experience,
many users build manually in their home directories out of habit.  At most sites, home
directories are remotely mounted volumes (e.g., NFS).  To compare Spack builds
with {\it typical} manual builds, we timed the same seven builds using
NFS.  Figure~\ref{fig:overhead} shows that building this way can be as much
as 62.7\% slower than using a temporary filesystem and
33\% slower on average. We therefore expect many users to notice
a speedup when Spack uses temporary space to build.

### 2015-07-31 implementation-packages.tex
Name change spack-paper-sc15 -> spack-sc15
Revisions and better overhead plot.

Figure~\ref{fig:mpileaks} shows the package for \mpileaks, a tool developed
at LLNL to find un-released MPI handle objects in parallel programs.
The {\tt MpiLeaks} class first provides simple metadata: a text description,
a homepage, and a download URL.
Next, two {\tt version} directives identify known versions and provide
MD5 checksums used to verify downloads.
Below this, two {\tt depends\_on}
directives indicate prerequisite packages that must be installed before \mpileaks.
Last, the {\tt install} method contains the commands used to build
the tool.  Spack's DSL allows shell commands to be invoked as Python functions,
and the {\tt install} method invokes {\tt configure},
{\tt make}, and {\tt make install} commands as a shell script would.

>>>>

Figure~\ref{fig:mpileaks} shows the package for \mpileaks, a tool developed
at LLNL to find un-released MPI handle objects in parallel programs.
The {\tt MpiLeaks} class provides simple metadata on lines 4-8: a text
description, a homepage, and a download URL.
Two {\tt version} directives on lines 10-11 identify known versions and provide
MD5 checksums to verify downloads.
The two {\tt depends\_on} directives on lines 13-14 indicate prerequisite
packages that must be installed before \mpileaks.
Last, the {\tt install} method on line 16 contains the commands for building.
Spack's DSL allows shell commands to be invoked as Python functions,
and the {\tt install} method invokes {\tt configure},
{\tt make}, and {\tt make install} as a shell script would.

### 2015-07-31 implementation-specs.tex
Revisions and better overhead plot.

Using the simple script in Figure~\ref{fig:mpileaks}, Spack can build many different
versions and configurations of the {\tt mpileaks} package.  In traditional port systems,
package code is structured to build a single version of a package, but in Spack, each
package file is a {\it template} that can be configured and built in many
different ways, according to a set of {\it parameters}.
Spack calls a single build configuration a {\it spec}. The {\tt spec} object
passed to the {\tt install} method is how Spack communicates dependencies and parameters to package authors.

\subsubsection{Structure}
To understand specs, consider the {\tt mpileaks} package structure.
Metadata in the {\tt Mpileaks} class ({\tt version}, {\tt depends\_on}, etc.) describe
its relationships with other packages.  There are two direct dependencies:
the {\tt callpath} library and {\tt mpi}.  Spack recursively inspects the class definitions
for each dependency and constructs a graph of their relationships.  The result
is a directed, acyclic graph (DAG).\footnote{Spack currently disallows
circular dependencies.}
%
To guarantee a consistent build and to avoid ABI compatibility issues, we construct
the DAG so that there is only {\it one} version of any package.  Note
the distinction: while Spack can install arbitrarily many configurations of any package,
no two configurations of the same package will ever appear in the same build DAG.

@@

Recall that Spack guarantees that there is only a single version of any package in
the spec DAG.  Within the same DAG, each dependency can be uniquely
identified by its package name alone.  Therefore, the user need not think about DAG
connectivity to add constraints.  She need only know that {\tt mpileaks}
somehow depends on {\tt callpath}.
For the same reason, the constraint order does not matter; dependency constraints
can appear in arbitrary order.

@@

So far, we have shown examples of specs used to request constraints from the
command line, when {\tt spack install} is invoked.  However, the user is not the only
source of constraints.  Applications may require specific versions of dependencies,
and it is often desirable to write constraints like this into a package file.  For
example, the ROSE compiler only builds with a certain version of the {\tt boost} library.
The {\tt depends\_on()} directives in Figure~\ref{fig:mpileaks} contain
simple package names.  However, a package name is also a spec, and
the same constraint syntax usable from the command line can be applied inside directives.
So, the ROSE compiler can simply write:
%
\begin{minted}[fontsize=\scriptsize]{python}
    depends_on('boost@1.54.0')
\end{minted}
%
This constraint will be incorporated into the initial DAG node generated from
the {\tt ROSE} package.  Dependencies can also be {\it conditional}.  For example,
the {\tt ROSE} package builds with different versions of boost, depending on the
compiler version. To handle this, directives also accept spec syntax as a
predicate in an optional {\tt when} parameter:
%
\begin{minted}[fontsize=\scriptsize]{python}
    depends_on('boost@1.47.0', when='%gcc@:4')
    depends_on('boost@1.54.0', when='%gcc@5:')
\end{minted}
%
The same notation is used to optionally build with libraries like MPI:
%
\begin{minted}[fontsize=\scriptsize]{python}
    depends_on('mpi', when='+mpi')
\end{minted}
%
and to ensure that specific patches are applied to Python source code
when it is built on Blue Gene/Q, with different compilers:
%
\begin{minted}[fontsize=\scriptsize]{python}
    patch('python-bgq-xlc.patch',   when='=bgq%xl')
    patch('python-bgq-clang.patch', when='=bgq%clang')
\end{minted}
%
Constraints in the {\tt when} clause are matched against the package spec.
Outside of directives, constraints can also be tested directly on the {\tt spec}
object in the {\tt install} method:
%
\begin{minted}[fontsize=\scriptsize]{python}
  def install(self, spec, prefix):
      if spec.satisfies('%gcc'):
          # Handle gcc
      elif spec.satisfies('%xl'):
          # Handle XL compilers
      ...
\end{minted}
%

>>>>

Using the simple script in Figure~\ref{fig:mpileaks}, Spack can build many different
versions and configurations of the {\tt mpileaks} package.  In traditional port systems,
package code is structured to build a single version of a package, but in Spack, each
package file is a {\it template} that can be configured and built in many
different ways, according to a set of {\it parameters}.
Spack calls a single build configuration a {\it spec}. 
Spack communicates dependencies and parameters to package authors using
the {\tt spec} argument of the {\tt install} method.

\subsubsection{Structure}
To understand specs, consider the {\tt mpileaks} package structure.
Metadata in the {\tt Mpileaks} class ({\tt version}, {\tt depends\_on}, etc.) describe
its relationships with other packages.  There are two direct dependencies:
the {\tt callpath} library and {\tt mpi}.  Spack recursively inspects the class definitions
for each dependency and constructs a graph of their relationships.  The result
is a directed, acyclic graph (DAG).\footnote{Spack currently disallows
circular dependencies.}
%
To guarantee a consistent build and to avoid 
Application Binary Interface (ABI) incompatibility, we construct
the DAG with only {\it one} version of each package.  Thus, while
Spack can install arbitrarily many configurations of any package,
no two configurations of the same package will ever appear in the same build DAG.

@@

Recall that Spack guarantees that there is only a single version of any package in
the spec DAG.  Within a spec, each dependency can be uniquely
identified by its package name alone.  Therefore, the user does not need to consider 
DAG connectivity to add constraints.  She need only know that {\tt mpileaks}
somehow depends on {\tt callpath}.
For the same reason, the constraint order does not matter; dependency constraints
can appear in arbitrary order.

@@

So far, we have shown examples of specs used to request constraints from the
command line, when {\tt spack install} is invoked.  However, the user is not the only
source of constraints.  Applications may require specific versions of dependencies,
and it is often desirable to write constraints like this into a package file.  For
example, the ROSE compiler~\cite{rose} only builds with a certain version of the {\tt boost} library.
The {\tt depends\_on()} directives in Figure~\ref{fig:mpileaks} contain
simple package names.  However, a package name is also a spec, and
the same constraint syntax usable from the command line can be applied inside directives.
So, the ROSE compiler can simply write:
%
\begin{minted}[fontsize=\scriptsize]{python}
    depends_on('boost@1.54.0')
\end{minted}
%
This constraint will be incorporated into the initial DAG node generated from
the {\tt ROSE} package.  Dependencies can also be {\it conditional}.  For example,
the {\tt ROSE} package builds with different versions of boost, depending on the
compiler version. To handle this, directives also accept spec syntax as a
predicate in an optional {\tt when} parameter:
%
\begin{minted}[fontsize=\scriptsize]{python}
    depends_on('boost@1.47.0', when='%gcc@:4')
    depends_on('boost@1.54.0', when='%gcc@5:')
\end{minted}
%
The same notation is used to optionally build with libraries like MPI:
%
\begin{minted}[fontsize=\scriptsize]{python}
    depends_on('mpi', when='+mpi')
\end{minted}
%
and to ensure that specific patches are applied to Python source code
when it is built on Blue Gene/Q, with different compilers:
%
\begin{minted}[fontsize=\scriptsize]{python}
    patch('python-bgq-xlc.patch',   when='=bgq%xl')
    patch('python-bgq-clang.patch', when='=bgq%clang')
\end{minted}
%
Constraints in the {\tt when} clause are matched against the package spec.
Outside of directives, constraints can also be tested directly on the {\tt spec}
object in the {\tt install} method:
%
\begin{minted}[fontsize=\scriptsize]{python}
  def install(self, spec, prefix):
      if spec.satisfies('%gcc'):
          # Handle gcc
      elif spec.satisfies('%xl'):
          # Handle XL compilers
      ...
\end{minted}
%

### 2015-07-31 implementation.tex
Revisions and better overhead plot.

\section{The Spack Package Manager}
\label{sec:implementation}
Based on our experiences at LLNL, we have developed
{\it Spack}, the Supercomputing PACKage manager.
Spack is written in Python.  We chose Python for its flexibility
and its wide use in the HPC community.
%
Like prior systems, Spack supports an arbitrary number of software
installations, and like Nix it can identify them with hashes.  Unlike any
prior system, Spack provides a language to specify and manage the
combinatorial space of HPC software configurations.

\noindent
Spack provides the following unique features:
\begin{enumerate}
\item Packages explicitly {\bf parameterized} by version, platform,
      compiler, options, and dependencies, to allow easy composition.
\item A novel, recursive {\bf spec syntax} for dependency graphs and constraints,
      which aids in managing the build parameter space.
\item {\bf Versioned virtual dependencies} to handle versioned, 
      ABI-incompatible interfaces like MPI.
\item A novel {\bf concretization} process that translates an abstract build
      specification into a full, concrete build specification.
\item A build environment with {\bf compiler wrappers} that enforce build
      consistency and simplify package writing.
\end{enumerate}

>>>>

\section{The Spack Package Manager}
\label{sec:implementation}
Based on our experiences at LLNL, we have developed
{\it Spack}, the \underline{S}upercomputing \underline{Pack}age manager.
Spack is written in Python, which we chose for its flexibility
and its increasing use in HPC.
%
Like prior systems, Spack supports an arbitrary number of software
installations, and like Nix it can identify them with hashes.  Unlike any
prior system, Spack provides a language to specify and manage the
combinatorial space of HPC software configurations.

\noindent
Spack provides the following unique features:
\begin{enumerate}
\item Composable packages, explicitly {\bf parameterized} by version, platform,
      compiler, options, and dependencies.
\item A novel, recursive {\bf spec syntax} for dependency graphs and constraints,
      which aids in managing the build parameter space.
\item {\bf Versioned virtual dependencies} to handle versioned, 
      ABI-incompatible interfaces like MPI.
\item A novel {\bf concretization} process that translates an abstract build
      specification into a full, concrete build specification.
\item A build environment that uses {\bf compiler wrappers} to enforce build
      consistency and simplify package writing.
\end{enumerate}

### 2015-07-31 introduction.tex
Revisions and better overhead plot.

This paper describes our experiences with the {\it Spack} package manager,
which we have developed at LLNL to manage increasing software complexity.
It makes the following contributions:
\begin{enumerate}
\item A novel, recursive syntax for concisely specifying and querying
      the large parameter space of HPC packages;
\item A build methodology that ensures packages find their dependencies
      regardless of users' environments;
\item Spack: An implementation of these concepts; and
\item Four use cases that detail LLNL's use of Spack to comcpose
      software configurations rapidly, to manage Python installations, and
      to implement site build policies, and to build an LLNL production code.
\end{enumerate}

>>>>

This paper describes our experiences with the {\it Spack} package manager,
which we have developed at LLNL to manage increasing software complexity.
It makes the following contributions:
\begin{enumerate}
\item A novel, recursive syntax for concisely specifying and querying
      the large parameter space of HPC packages;
\item A build methodology that ensures packages find their dependencies
      regardless of users' environments;
\item Spack: An implementation of these concepts; and
\item Four use cases detailing LLNL's use of Spack to compose
      software configurations rapidly, to manage Python installations,
      to implement site build policies, and to build an LLNL production
      code with 47 dependencies in 32 configurations.
\end{enumerate}

### 2015-07-31 limitations.tex
Revisions and better overhead plot.

Spack can require more disk space than a module-based system, as otherwise 
identical packages with different dependencies must be built separately. In 
our view, the significant reduction in complexity for the end user justifies 
this cost.

The use of Python has been a barrier for some users, and
a major attraction for others. For those who would otherwise create local
workarounds, Python's flexibility has allowed them to extend local
Spack packages to suit their needs. Such customizations frequently
guide our priorities for new core features. The
benefits of using a popular, rapidly emerging HPC language for our
embedded DSL outweigh any costs of the learning curve.

In the near term, we are actively working to grow a community around Spack 
and to build a larger base of contributors. On the technical side, we will
increase build robustness by deploying a continuous testing system
at LLNL.  Finally, to support the growing number of
HPC languages, runtimes, and programming models, we will add
capabilities to Spack that allow packages to depend on particular
compiler features.  More and more, our codes are beginning to rely
on advanced compiler capabilities, like {\tt C++11} language features,
OpenMP versions, and GPU compute capabilities. Ideally, Spack
will find suitable compilers and ensure ABI consistency when many
such features are in use.

>>>>

Spack requires more disk space than a module-based system, as otherwise 
identical packages with different dependencies must be built separately. 
The exact space overhead depends on the structure of the installed
software; some builds can share more dependency libraries than others
(see Figure~\ref{fig:reuse}).
In our view, the significant reduction in complexity for the end user
justifies this cost.

The use of Python has been a barrier for some users, and
a major attraction for others. For those who would otherwise create local
workarounds, Python's flexibility has allowed them to extend local
Spack packages to suit their needs. Such customizations frequently
guide our priorities for new core features. Some users have found the 
learning curve to be too steep, but this only prevents them from 
packaging software. They still use Spack as a command-line tool.

In the near term, we are actively working to grow a community around Spack 
and to build a larger base of contributors. We plan to
increase build robustness by deploying a continuous testing system
at LLNL.  Finally, to support the growing number of
HPC languages, runtimes, and programming models, we will add
capabilities to Spack that allow packages to depend on particular
compiler features.  More and more, our codes are relying
on advanced compiler capabilities, like {\tt C++11} language features,
OpenMP versions, and GPU compute capabilities. Ideally, Spack
will find suitable compilers and ensure ABI consistency when many
such features are in use.

### 2015-07-31 usecase-python.tex
Name change spack-paper-sc15 -> spack-sc15
Revisions and better overhead plot.

Python is becoming increasingly popular for HPC applications, 
due to its flexibility as a language and its excellent support
for calling into fast, compiled numerical libraries.
Python is an interpreted language, but one can use it
as a friendlier interface to compiled libraries like FFTW, ATLAS, and
other linear algebra libraries.  Many LLNL code teams use Python in this manner.
%
LC supports Python installations for several application teams,
and maintaining these repositories has grown increasingly complex over
time. The problems are similar to those that drove us to create
Spack: different teams want different Python libraries with different
configurations.

Existing Python package managers are either language-specific~\cite{eby:setuptools}, 
or they do not support building from source~\cite{anaconda,conda}. None
handles multi-configuration builds.  More glaringly, Python extensions
are usually installed into the Python interpreter's prefix.
This makes it impossible to install multiple versions.\footnote{{\tt setuptools}
has support for multiple versions via {\tt pkg_resources}, 
but this requires modifications to client code.}
Installing each extension in its own prefix enables combinatorial versioning,
but it requires users to add packages to the {\tt PYTHONPATH} variable at runtime.

Per Spack's design philosophy, we wanted a way to easily manage many different
versions, but {\it also} to provide a baseline set of extensions {\it without}
requiring environment settings.
%
To support this mode of operation, we added the concept of {\tt extension} packages
to Spack. Python modules use the {\tt extends('python')} directive instead of
{\tt depends\_on('python')}.
Each module installs into its own prefix like any other package,
and each module depends on a particular Python installation.
But extensions can be {\tt activated} or {\tt deactivated} an
within the dependent Python installation.  The {\tt activate} operation
symbolically links each file in the extension prefix into the Python
installation prefix, as though it were installed directly. If any file
conflict would arise from this operation, {\tt activate} fails.
Similarly, the {\tt deactivate} operation removes the symbolic links and restores
the Python installation to its  pristine state.

>>>>

Python is becoming increasingly popular for HPC applications,
due to its flexibility as a language and its excellent support
for calling into fast, compiled numerical libraries.
Python is an interpreted language, but one can use it
as a friendlier interface to compiled libraries like FFTW, ATLAS, and
other linear algebra libraries.  Many LLNL code teams use Python in this manner.
%
LC supports Python installations for several application teams.
Maintaining these repositories has grown increasingly complex over
time. The problems are similar to those that drove us to create
Spack: different teams want different Python libraries with different
configurations.

Existing Python package managers either do not support building from
source~\cite{anaconda,conda}, or they are language-specific~\cite{eby:setuptools}.
None handles multi-configuration builds.  More glaringly, Python extensions
are usually installed into the Python interpreter's prefix.
This makes it impossible to install multiple versions.\footnote{{\tt setuptools}
has support for multiple versions via {\tt pkg_resources},
but this requires modifications to client code.}
Installing each extension in its own prefix enables combinatorial versioning,
but it requires users to add packages to the {\tt PYTHONPATH} variable at runtime.

Per Spack's design philosophy, we wanted a way to easily manage many different
versions, but {\it also} to provide a baseline set of extensions {\it without}
requiring environment settings.
%
To support this mode of operation, we added the concept of {\tt extension} packages
to Spack. Python modules use the {\tt extends('python')} directive instead of
{\tt depends\_on('python')}.
Each module installs into its own prefix like any other package,
and each module depends on a particular Python installation.
But extensions can be {\tt activated} or {\tt deactivated}
within the dependent Python installation.  The {\tt activate} operation
symbolically links each file in the extension prefix into the Python
installation prefix, as though it were installed directly. If any file
conflict would arise from this operation, {\tt activate} fails.
Similarly, the {\tt deactivate} operation removes the symbolic links and restores
the Python installation to its  pristine state.

### 2015-07-31 usecases.tex
edits to usecases
Change name of tex file to 'spack-paper-sc15'
combinatoric -> combinatorial
Add ARES dependency figure.
Name change spack-paper-sc15 -> spack-sc15
Revisions and better overhead plot.

We have recently using Spack in production at 
Livermore Computing.  We have already outlined a number of the
experiences that drove us to develop Spack.
In this section, we describe real-world use cases that Spack
has been used to address.  In some cases, Spack had to be 
adapted to meet production needs, and we describe how its
flexibility has allowed us to put together solutions quickly.

\input{usecase-combinatoric}
\input{usecase-python}
\input{usecase-policy}

>>>>

We have recently begun using Spack in production at
Livermore Computing (LC).  We have already outlined a number of the
experiences that drove us to develop Spack.
In this section, we describe real-world use cases that Spack
has addressed.  In some cases, Spack had to be
adapted to meet production needs, and we describe how its
flexibility has allowed us to put together solutions quickly.

\input{usecase-combinatorial}
\input{usecase-python}
\input{usecase-policy}
\input{usecase-ares}
